---
title: "The Blue Zones: Why do some areas of the world have higher life expectancies than others?"
author: "Group 9"
date: "2022-11-22"
output:
  html_document:
    theme: journal
    highlight: tango
    number_sections: yes
    toc: yes
    toc_float: no
  pdf_document:
    toc: yes
editor_options: 
  markdown: 
    wrap: sentence
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load packages}
library(vroom)
library(tidyverse)
library(lubridate)
library(ggmap)
library(ggrepel)
library(gridExtra)
library(pander)
library(here)
library(sf)
library(janitor)
library(skimr)
library(leaflet)
library(tmap)
library(tmaptools)
library(hrbrthemes)
library(mapview)
library(viridis)
library(dplyr)
```


## (1) A statement of the question or purpose. What problems or questions did you set out to analyse? What were the key issues raised? 

(1) Our goal is to analyze different factors, such as; nutrition, lifestyle, exercise, climate, genetics, hygiene,... and the impact they have on longevity in different parts of the world.

More specifically we aim to answer these questions:

a. Why do people living in blue zones live longer than others ?
b. How and why did longevity change over time for these countries ?
c. Which factors have the biggest influence on life expectancy ?


## (2) The background and preparation for conducting the project. How did you prepare for the project? What sources or background readings did you consult? What information did you use in developing your ideas from the conceptual stage to the finished project? 

We read through some articles about Blue Zones. Blue zones are regions in the world where people are speculated to live longer than average; the term was coined by Gianni Pes, Michel Poulain and Dan Buettner. There are five blues zones: Okinawa (Japan), Sardinia (Italy), Nicoya (Costa Rica), Icaria (Greece) and Loma Linda (California). We also looked up some research about some potential factors that could contribute to life expectancy in a region (eg. infant mortality rate, GDP, education level, and different deseases)


## (3) Methodology. What did you do, and how did you do it? What statistical techniques did you use — for instance, scatterplots, correlation, confidence intervals, linear/logistic regression? 

We used hypothesis testing to check correlations between factors affecting life expectancy and life expectancy. We used linear regressions to identify  variables that are significantly impacting life expectancy in a particular country in different time points, and then we adjusted the regression model based on the initial regression result to increase its explaining power. We also used cluster analysis based on GDP and life expectancy to cluster world nations.

## (4) Results and conclusions. This is where you summarise and present your data analyses and communicate your main results. What did you find out? This might include tables, graphs, or verbal summaries. What did you learn about the problem or question you set out to investigate? 


#Data Cleaning

```{r infant-inmortality}


df1 <- read_csv(file = 'data viz dataset/(NEEDS CLEAN)-infant-inmortality.csv',skip=4, show_col_types = FALSE) 
df1 <-df1[,-67]
df1 <-df1[,-66]

df1 <- df1 %>% 
  drop_na() %>% 
  pivot_longer(names_to = "year",
               values_to = "infant_inmortality",
               cols = -(1:4)) %>% 
  janitor::clean_names()


glimpse(df1)

write.table(df1, file = "data viz dataset/CLEANED-infant-mortality.csv", row.names = FALSE, na = "", sep = ",")
#storing the data in csv file, removing row name and column name

```


```{r life-expectancy}

#read in data, remove the first 3 rows
df2 <- read_csv(file = 'data viz dataset/(NEEDS CLEAN) life-expectancy-2020.csv',skip=4, show_col_types = FALSE) 
df2 <-df2[,-67]
df2 <-df2[,-66]

df2 <- df2 %>% 
  drop_na() %>% 
  pivot_longer(names_to = "year",
               values_to = "life_expectancy",
               cols = -(1:4)) %>% 
  janitor::clean_names()


glimpse(df2)


write_csv(df2, file = "data viz dataset/CLEANED-life-expectancy-2020.csv", na = "")
#storing the data in csv file, removing row name and column name


```



## Regression on life-expectancy/potential factor dataset

#preparing data for regression
```{r}


lc <- read.csv('data viz dataset/CLEANED-Life Expectancy - potential factors.csv') %>% 
  clean_names()

lc_clean <- na.omit(lc)


#filtering data to the latest available year (2014)
lc_2014=lc_clean%>%filter(year==2014)

lc_2004=lc_clean%>%filter(year==2004)

lc_2000=lc_clean%>%filter(year==2000)

```

#linear model on year 2014
```{r}


lm1 <-lm(life_expectancy ~ bmi + polio + percentage_expenditure + hepatitis_b +measles +total_expenditure + diphtheria + hiv_aids + gdp + income_composition_of_resources + schooling+ alcohol,lc_2014)

summary(lm1)

#adjusting the linear model 

lm_new1=lm(life_expectancy ~  total_expenditure + hiv_aids + income_composition_of_resources, lc_2014)
  
summary(lm_new1)
  
```


Based on multi-variable linear regression result, we can see that, in 2014, Income Composition of Resources, and HIV rate significantly impacts the average life expectancy in a particular country. （We left out Adult Mortality to avoid perfect fit problem)


#linear model on year 2004
```{r}


lm2 <-lm(life_expectancy ~ bmi + polio + percentage_expenditure + hepatitis_b +measles +total_expenditure + diphtheria + hiv_aids + gdp + income_composition_of_resources + schooling+ alcohol,lc_2004)


summary(lm2)

#adjusting the linear model 

lm_new2=lm(life_expectancy ~ bmi + hiv_aids + income_composition_of_resources +gdp +alcohol+schooling,lc_2004)
  
summary(lm_new2)

```

However, when we run multi-variable linear regression on year 2004 dataset, we can see that, in 2004, Income Composition of Resources no longer has an  impact on life expectancy. In its place, Schooling (education level), HIV rate, and under-five deaths are significantly impacting average life expectancy in a country.


#linear model on year 2000
```{r}


lm3 <- lm(life_expectancy ~ bmi + polio + percentage_expenditure + hepatitis_b +measles +total_expenditure + diphtheria + hiv_aids + gdp + income_composition_of_resources + schooling+ alcohol,lc_2000)

summary(lm3)

lm_new3=lm(life_expectancy ~ hepatitis_b  + adult_mortality + hiv_aids +schooling + hepatitis_b,lc_2000)
  
summary(lm_new3)

```

Based on multi-variable linear regression result, we can see that, in 2000, Schooling, infant (under-5 death rate) and HIV rate significantly impact the average life expectancy in a particular country.



##Regression Visualization 


```{r some graph}

library(gapminder)
gapminder %>% ggplot(aes(x = continent, y=lifeExp, group_by = year, colour = continent))+
  facet_wrap(~year)+
  geom_boxplot()+
  labs(title = "Life Expectancy of Different Continents Over the Years")+
  theme_bw()+
  theme(legend.position = "None")

```


Before doing the actual analytics, we assumed that GDP would be correlated with life-expectancy, but the linear regression result suggested otherwise. Therefore, we did further clustering analysis on GDP and life-expectancy and tried to identify the reasons behind why GDP and life-expectancy is not linearly correlated.

```{r clustering gdp lifeexp}

library(factoextra)

lifeExp_factors <- read_csv(file = "data viz dataset/CLEANED-Life Expectancy - potential factors.csv",show_col_types = FALSE) 

lifeExp_factors_cluster <- lifeExp_factors %>% 
  clean_names() %>% 
  filter(year==2014) %>% 
  select("life_expectancy", "gdp") %>% 
  na.omit()

# lifeExp_factors %>% 
#   clean_names() %>% 
#   filter(year==2014) %>% 
#   select("country", "life_expectancy", "gdp") %>% 
#   slice_max(gdp, n=10) %>% 
#   arrange(desc(gdp))
# 
# lifeExp_factors %>% 
#   clean_names() %>% 
#   filter(year==2014) %>% 
#   select("country", "life_expectancy", "gdp") %>% 
#   slice_max(life_expectancy, n=10) %>% 
#   arrange(desc(life_expectancy))

model_kmeans <- eclust(lifeExp_factors_cluster, "kmeans", k=3, graph = TRUE)

#Let's check the components of this object.
summary(model_kmeans)


library(ggpubr)

lifeExp_factors_withClusters<-mutate(lifeExp_factors_cluster, cluster = as.factor(model_kmeans$cluster))

ggplot(lifeExp_factors_withClusters, aes(x = gdp, y = life_expectancy, color =  as.factor(cluster))) +
  geom_jitter()+
  labs(color = "Cluster",
       title = "Clustering Countries by GDP and Life Expectancy")

```


### (5) Discussion and critique. What did you learn about the process of carrying out your project? What went wrong, and how could you improve it next time? For instance, did any sources of bias creep into your survey or experiment? What advice would you give future students? 

(1)At first, we used every variable in the data set to do the multi-variable regression, without assessing potential perfect fit and co-linearity problem. For example, "Schooling" and "Income" are likely to be linearly related; "Infant mortality" and "under-five deaths" are likely to be linearly relater; and "Adult mortality" is likely to perfectly explain the dependent variable "life expectancy". Therefore, we took these variables out of the model. 

(2)For features like "infant mortality", and "under-five deaths", even if they are significantly related to the dependent variable, it cannot generate any useful explanatory result. There fore we should focus on variables that could potentially be well explained.


For future references, we should be more careful when dealing with feature choice and feature engineering, especially focus on variable's explaining power and multi-colinearity problem.


